---
title: "data_wrangling2"
author: "Xuehan Yang"
date: "2021/10/18"
output: github_document
---
Websties are composed of html + css. Content is html, css controls the appearance. So we can use CSS selector to select unique identifier of specific html content.

download html using read_html,
extract elements with html_elements() and CSS selector
extract contents with html_text() and html_table()

APIs

provide a way to communicate with software. request specific data from a server. http

httr contains a collection of tools for contructing HTTP requests.
focus on GET
Query

API data formats
sometimes, go to the web and click download. # But data changes often. this way is not reproducible.
Most cases, get JavaScriptObjectNotation(JSON) JSON files can be parsed in R using jsonlite.


```{r}
library(tidyverse)
library(rvest)
library(httr)
```


# NSDUH

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url) # this step is done on the Internet

drug_use_html
```

```{r}
drug_use_html %>% 
  html_table()
```

get the contents from the first list element
```{r}
table_marj =
  drug_use_html %>% 
  html_table() %>% 
  first() # if you want to take 2-n tables, use nth(2)
```

```{r}
table_marj =
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
  slice(-1) # remove the first row which is a note.
```

# Star Wars

Get some star wars data...

```{r}
cost = read_html("https://www.bestplaces.net/cost_of_living/city/new_york/new_york") %>% 
  html_table(header = TRUE) %>% 
  first()
```

CSS Selectors
```{r}
sw_url = "https://www.imdb.com/list/ls070150896/"

swm_html = 
  read_html(sw_url)
```


```{r}
title_vec = 
  swm_html %>%
  html_elements(".lister-item-header a") %>%
  html_text()

gross_rev_vec = 
  swm_html %>%
  html_elements(".text-small:nth-child(7) span:nth-child(5)") %>%
  html_text()

runtime_vec = 
  swm_html %>%
  html_elements(".runtime") %>%
  html_text()

swm_df = 
  tibble(
    title = title_vec,
    rev = gross_rev_vec,
    runtime = runtime_vec)
```

## Napoleon

Dynamite dynamite

```{r}
dynamite_url  = "https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber=1"

dynamite_df = read_html(dynamite_url)

dy_review_title =
  dynamite_df %>% 
  html_elements(".a-text-bold span") %>% 
  html_text()

dy_df = tibble(
  title = dy_review_title
)

```

## Using an API

Get some data from API about water

```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")
```

Let's see what JSON looks like
```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>%
  jsonlite::fromJSON() %>%
  as_tibble()
```

BRFSS data via API
```{r}
brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) %>% #api dont just send you everyting, it is kind of protection.
  content("parsed")
```
NYC restaurant project for reference.

What about Poke!
```{r}
poke = 
  GET("http://pokeapi.co/api/v2/pokemon/1") %>%
  content()

poke$name

poke$height
```




